A Fast kNN Algorithm Using Multiple Space-Filling Curves

### 2022-04-15 2022/Sensors/kNN.tex

Indeed, the point $x \in [0,1]$ on the one-dimensional scale has only left and right neighbours, whereas its corresponding object $y(x) \in R^N$ can have neighbours along $N$ coordinate directions. As a result, when using Peano curve type mappings, close objects $y^\prime$, $y^{\prime\prime}$ in the $N$-dimensional space can have their quite distant corresponding preimages $x^\prime$, $x^{\prime\prime}$ on the interval $[0,1]$. In Fig. 1 the green dots show two objects that are close in the two-dimensional space, while their preimages on the one-dimensional scale are far away from each other. The blue dots correspond to two objects, the distance between which is retained when changing to the one-dimensional scale.

\begin{figure}
\center
\includegraphics[width=0.5\linewidth]{fig1.png}
\caption{Peano curve}
\label{fig1}
\end{figure}

@@

Any subcube of the $M$-th partition of the hypercube $D_0$ will have an edge of length $2^{-(M-1)}$ and will be denoted by $D_0(M, v)$, where $v$ is the left boundary point of the subinterval $d(M, v)$ corresponding to this subcube. Then the evolvents $y^l(x)=\{y_1^l(x), ..., y_N^l(x)\}$, whose coordinates are determined by the conditions
\begin{equation}\label{maps}
y_i^l(x) = y_i^{l-1}(x) + 2^{-l}, \; 1 \leq i \leq N, \; 1 \leq l \leq L,
\end{equation} 
map the interval $[0,1]$ onto the corresponding hypercubes  $D_l, \; 1 \leq l \leq L$ (the broken line in Fig.~\ref{fig2} shows the image of the interval $[0,1]$, which is obtained by using the evolvent $y^0(x), x \in [0,1]$).

@@

Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Fig.~\ref{fig2}), then, by introducing an additional function 
\[
g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
\] 
the original hypercube $D$ can be represented as
\[
D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
\] 
i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
\[
y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
\]

@@

As an illustration, Fig. \ref{fig2} shows a family of space-filling curves in extended domains $D_l$, where a square with a dark border highlights the hypercube $D$ belonging to each extended domain. The dots indicate proximate objects in the domain $D$, the distance between which will vary depending on the one-dimensional scale used. The effect of maintaining distance proximity on one of the scales is clearly visible. For example, objects marked with green dots will be far away on the first two one-dimensional scales and close to each other on the third one. Objects marked in blue will be far away on the first and third scales and close to each other on the second scale.

@@

The running times used by the different methods (kNN-ME, procedure A, kNN-ME, procedure B, kNN-KD) for DS-Random-1 and DS-Random-2 datasets with $N = 2$ are shown in Fig. \ref{fig3}. The abscissa axis represents the number of nearest neighbours that the algorithm is searching for, the ordinate axis represents the running time of the method.

@@

Running times of the methods for nearest neighbours search with $N = 3, 4$ are given in Figs. \ref{fig4}, \ref{fig5} respectively.

@@

The time required to find $K$ nearest neighbours on the CarEvoluation dataset is shown in Fig.~\ref{fig6a}. In this experiment, the number of evolvents used simultaneously was 7. 
\begin{figure}
\widefigure
\includegraphics[width=0.5\linewidth]{fig6a.pdf}
\caption{Comparison of neighbour search times for the CarEvoluation dataset\label{fig6a}}
\end{figure}
The results of the comparison on this dataset also demonstrate that the kNN-ME method is faster compared to the kNN-KD method.

>>>>

Indeed, the point $x \in [0,1]$ on the one-dimensional scale has only left and right neighbours, whereas its corresponding object $y(x) \in R^N$ can have neighbours along $N$ coordinate directions. As a result, when using Peano curve type mappings, close objects $y^\prime$, $y^{\prime\prime}$ in the $N$-dimensional space can have their quite distant corresponding preimages $x^\prime$, $x^{\prime\prime}$ on the interval $[0,1]$. In Figure~\ref{fig:Peano} the green dots show two objects that are close in the two-dimensional space, while their preimages on the one-dimensional scale are far away from each other. The blue dots correspond to two objects, the distance between which is retained when changing to the one-dimensional scale.

\begin{figure}
\center
\includegraphics[width=0.5\linewidth]{fig1.png}
\caption{Peano curve}\label{fig:Peano}
\label{fig1}
\end{figure}

@@

Any subcube of the $M$-th partition of the hypercube $D_0$ will have an edge of length $2^{-(M-1)}$ and will be denoted by $D_0(M, v)$, where $v$ is the left boundary point of the subinterval $d(M, v)$ corresponding to this subcube. Then the evolvents $y^l(x)=\{y_1^l(x), ..., y_N^l(x)\}$, whose coordinates are determined by the conditions
\begin{equation}\label{maps}
y_i^l(x) = y_i^{l-1}(x) + 2^{-l}, \; 1 \leq i \leq N, \; 1 \leq l \leq L,
\end{equation} 
map the interval $[0,1]$ onto the corresponding hypercubes  $D_l, \; 1 \leq l \leq L$ (the broken line in Figure~\ref{fig2} shows the image of the interval $[0,1]$, which is obtained by using the evolvent $y^0(x), x \in [0,1]$).

@@

Since the hypercube $D$ from (\ref{D}) belongs to the common part of the family of hypercubes (\ref{hypercubes}) (the boundary of the hypercube $D$ is highlighted in Figure~\ref{fig2}), then, by introducing an additional function 
\[
g_0(y) = \max \{ |y_i|-2^{-1}: \; 1 \leq i \leq N \},
\] 
the original hypercube $D$ can be represented as
\[
D = \{y^l(x): x \in [0,1], \; g_0(y^l(x)) \leq 0 \}, \; 0 \leq l \leq L,
\] 
i.e. $g_0(y) \leq 0$, if $y \in D$, and $g_0(y) > 0$ if otherwise. Hence, any point $y \in D$ has its preimage $x^l \in [0,1]$ at every mapping $y^l(x), \; 0 \leq l \leq L$, i.e.
\[
y = y^l(x^l), \; x^l \in [0, 1], \; 0 \leq l \leq L.
\]

@@

As an illustration, Figure~\ref{fig2} shows a family of space-filling curves in extended domains $D_l$, where a square with a dark border highlights the hypercube $D$ belonging to each extended domain. The dots indicate proximate objects in the domain $D$, the distance between which will vary depending on the one-dimensional scale used. The effect of maintaining distance proximity on one of the scales is clearly visible. For example, objects marked with green dots will be far away on the first two one-dimensional scales and close to each other on the third one. Objects marked in blue will be far away on the first and third scales and close to each other on the second scale.

@@

The running times used by the different methods (kNN-ME, procedure A, kNN-ME, procedure B, kNN-KD) for DS-Random-1 and DS-Random-2 datasets with $N = 2$ are shown in Figure~\ref{fig3}. The abscissa axis represents the number of nearest neighbours that the algorithm is searching for, the ordinate axis represents the running time of the method.

@@

Running times of the methods for nearest neighbours search with $N = 3, 4$ are given in Figures \ref{fig4}, \ref{fig5} respectively.

@@

The time required to find $K$ nearest neighbours on the CarEvoluation dataset is shown in Figure~\ref{fig6a}. In this experiment, the number of evolvents used simultaneously was 7. 
\begin{figure}
\widefigure
\includegraphics[width=0.5\linewidth]{fig6a.pdf}
\caption{Comparison of neighbour search times for the CarEvoluation dataset\label{fig6a}}
\end{figure}
The results of the comparison on this dataset also demonstrate that the kNN-ME method is faster compared to the kNN-KD method.
